---
title: "STA 104: Take Home Project  \nLooking into States that are being affected from Covid-19"
author: "Aman Singh & Peter Mueller  \namasin@ucdavis.edu  \npetermueller@ucdavis.edu  \nUC Davis  \n Instructor: Amy T. Kim  "
date: "2/28/2021"
output: pdf_document
header-includes:
  - \usepackage{titling}
  - \usepackage{tikz}
  - \pretitle{\begin{center}}
  -  \posttitle{\end{center}}
  -  \usepackage{blindtext}
---
\newpage
```{r, echo = F}
#cuttingoffcode
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=70),tidy=TRUE)
```
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

```{r,warning=FALSE,echo=FALSE,results=FALSE,include=FALSE}
#importing dataset
library(readr)
CovidA <- read_csv("CovidA.csv")
```

\circled{I} \textbf{Introduction}

\circled{II} \textbf{Summary of Data}
Thinking of doing the analysis on California, New York City, Florida, and Hawaii.
subsetting the data:

```{r,warning=FALSE,echo=FALSE,results=FALSE,include=FALSE}
#subsetting data
CovidA_sub = subset(CovidA,CovidA$State == "California"|CovidA$State == "New York City" | CovidA$State == "Florida" | CovidA$State == "Hawaii")
```

\circled{III} \textbf{Analysis}
```{r,warning=FALSE,echo=FALSE,results=FALSE,include=FALSE}
library(ggplot2)
library(dplyr)
```
```{r,warning=FALSE,echo=FALSE}
ggplot(data = CovidA_sub, aes(Month, Death)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color="steelblue") + 
  labs(title = "Monthly Covid Deaths compared to States",
       subtitle = "Analysis of California, Florida, Hawaii & New York City Covid-19 Deaths ",
       y = "Deaths", x = "Month") + 
  facet_wrap(~ State)
```


\circled{IV} \textbf{Interpretation}

\circled{V} \textbf{Conclusion}

\newpage
\circled{VI} \center{\textbf{Code Appendix}}
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```
